---
title: "Automating linear regression"
author: "Dan"
date: "2022-10-27"
output: html_document
---

```{r load libraries}
library(dplyr) # for data munging 
library(foreign) # we'll need to import the .SAV file into R
library(ggplot2) # visualization : better aesthetic add-ons to effect plots
library(effects) # for plotting the marginal effects of variables 
#library(olsrr) # for model diagnostics 
library(tidyr)
library(cowplot)
library(patchwork)
library(DT)
library(ggeffects)
```

Import the data into R 

```{r data import}
mtcarss <- mtcars
```

Run a logistic model on the data with the response variable being contraceptive_use.  

```{r data prepping }
mtcarss$vs <- factor(mtcarss$vs)
mtcarss$am <- factor(ifelse(mtcarss$am == 0, "automatic", "manual"))
mtcarss$gear <- factor(mtcarss$gear)
mtcarss$carb <- factor(mtcarss$carb)

row.names(mtcarss) <- NULL

```

# Automation  
We can make a function that takes in two parameters, the response variable and the data, runs the model and checks if the p-values are significant or not then prints a list of two items, significant variables and those that are not.  

```{r}

# define the response variable 
resvar <- "mpg"
mpg <- mtcarss$mpg

reg_auto_model <- function(dat,rvar){
  
  res_var <- eval(as.name(rvar))
  ndat <- dplyr::select(dat, -all_of(rvar))

  regModel <- lm(res_var ~ ., data=ndat)
  #reglogmodel <- lm(log(mpg) ~ ., data = dat)
  
  return(list(summary(regModel),coef(summary(regModel))[,'Pr(>|t|)'],regModel,reglogmodel))
  
}

regressionModel <- reg_auto_model(dat=mtcarss, rvar = resvar) 
regressionModel[[2]]

```


# Evaluating the variables to flag those that are significant and otherwise 

```{r}

regpvalues <- regressionModel[[2]] %>% as.data.frame()
regpvalues$variable <- rownames(regpvalues)
names(regpvalues)[1] <- "pval"
rownames(regpvalues) <- NULL

# The following table highlights variables which are significant and those that are not
regpvalues <- regpvalues %>% mutate(significance = case_when(
  pval < 0.5 ~ "significant",
  pval >= 0.5 ~ "Insignificant"
))

datatable(regpvalues)

```

# Visualizing the effect of variables on the outcome 
```{r}
regmodel.out <- regressionModel[[3]]
regpredictors <- colnames(mtcarss)
regpredictors <- regpredictors[regpredictors != resvar]

do.call(patchwork::wrap_plots, lapply(regpredictors, function(x) {
  print(plot(ggpredict(regmodel.out,eval(x)), show.title=FALSE, colors="bw"))
})) -> regplot

regplot

```

# Model diagnostics  

Regression assumptions

Linear regression makes several assumptions about the data, such as :

    Linearity of the data. The relationship between the predictor (x) and the outcome (y) is assumed to be linear.
    Normality of residuals. The residual errors are assumed to be normally distributed.
    Homogeneity of residuals variance. The residuals are assumed to have a constant variance (homoscedasticity)
    Independence of residuals error terms.

You should check whether or not these assumptions hold true. Potential problems include:

    Non-linearity of the outcome - predictor relationships
    Heteroscedasticity: Non-constant variance of error terms.
    Presence of influential values in the data that can be:
        Outliers: extreme values in the outcome (y) variable
        High-leverage points: extreme values in the predictors (x) variable

All these assumptions and potential problems can be checked by producing some diagnostic plots visualizing the residual errors.

```{r}

par(mfrow = c(2, 2))
plot(regmodel.out)

```

The diagnostic plots show residuals in four different ways:

    Residuals vs Fitted. Used to check the linear relationship assumptions. A horizontal line, without distinct patterns is an indication for a linear relationship, what is good.

    Normal Q-Q. Used to examine whether the residuals are normally distributed. It’s good if residuals points follow the straight dashed line.

    Scale-Location (or Spread-Location). Used to check the homogeneity of variance of the residuals (homoscedasticity). Horizontal line with equally spread points is a good indication of homoscedasticity. 

    Residuals vs Leverage. Used to identify influential cases, that is extreme values that might influence the regression results when included or excluded from the analysis. 

